Worflow

### 2017-06-07 R/workflow.Rmd

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
#bioconductor:
library(scRNAseq)
library(scone)
library(clusterExperiment)
#github
library(zinbwave)
library(slingshot)
#CRAN
library(ggplot2)
library(gplots)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(RColorBrewer)
library(digest)
library(rARPACK)
set.seed(20)
if(packageVersion("clusterExperiment")<'1.3.0.9006') stop("must have current develop version to avoid bugs")
```

```{r}
runZinb <- T
runClus <- T
NCORES <- 10
```

@@

EAP: Fanny, we should really be using the RSEC function for all the steps of the clustering. I've put in the code, but I set `eval=FALSE` because I don't have the data to test it. I also am not sure that we want to set the options for `subsampleArgs` and `seqArgs` since those are suppose to be sort of esoteric concerns...

@@

```{r rsec_50}
W <- colData(se)[, grepl('^W', colnames(colData(se)))]
W <- as.matrix(W)
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
  ceObj <- RSEC(t(W), k0s = 4:15, alphas = c(0.1), betas = 0.8,
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = 7,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none")
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

```{r examineMergeClusters,eval=FALSE}
#re-does merging simpling to make plot 
mergeClusters(ceObj, mergeMethod = "locfdr",
              plotInfo = "mergeMethod", cutoff = 0.01)
```

@@

```{r}
sampleData <- data.frame(ours = primaryCluster(ceObj))
plotHeatmap(assays(se)$normalizedValues,
            main = 'Normalized values, 1000 most variable genes',
            clusterSamplesData = ceObj@dendro_samples,
            sampleData = as.matrix(sampleData))
```

@@

>>>>

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
#bioconductor:
library(scRNAseq)
library(scone)
library(clusterExperiment)
#github
library(zinbwave)
library(slingshot)
#CRAN
library(ggplot2)
library(gplots)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(RColorBrewer)
library(digest)
library(rARPACK)
set.seed(20)
if(packageVersion("clusterExperiment")<'1.3.0.9008') stop("must have current develop version to avoid bugs")
```

```{r}
runZinb <- T
runClus <- T
NCORES <- 7
```

@@

@@

EP: I updated it to work on a SE object so that it has the meta data. If you have a summarized experiment object with W already, you could use that as long as assay(seObj) gives W.

```{r rsec_50}
W <- colData(se)[, grepl('^W', colnames(colData(se)))]
W <- as.matrix(W)
fn <- '../data/RSEC_W.rda'
if (runClus & !file.exists(fn)){
	metadata$clusterLabels[is.na(metadata$clusterLabels)]<- -2 #symbol for samples missing from original clustering
  seObj<-SummarizedExperiment(t(W),colData=metadata)
 print(system.time(  ceObj <- RSEC(seObj, k0s = 4:15, alphas = c(0.1), betas = 0.8,
                clusterFunction = "hierarchical01", minSizes=1,
                ncores = NCORES, isCount=FALSE,
                subsampleArgs = list(resamp.num=100,
                                     clusterFunction="kmeans",
                                     clusterArgs=list(nstart=10)),
                seqArgs = list(k.min=3, top.can=5), verbose=TRUE,
                combineProportion = 0.7,
                mergeMethod = "none")
				))
  save(ceObj, file = fn)
}else{
  load(fn)
}
```

@@

EP: I don't think the merging step on the W makes a whole lot of sense -- the method is irrelevant. The merging is based on calculating the % of genes found significant (the specific method is arbitrary). The best thing would be to replace the W with residuals in the assay of `ceObj` (or whatever data that you will do the DE on for the time stuff below), and then run the merging step on that data.  I'm not particularly fond of `locfdr`. It was probably the method that gave the best merging to Russell and Diya. You'd really have to run `mergeClusters` setting `plotInfo="all"` and look at the results and decide both the cutoff level and the method.

EP: Also, if you don't save the output of `mergeClusters` it doesn't update `ceObj`. I was calling it for just the resulting plots, since it was already merged in RSEC above. I've changed to code to update ceObj below.

```{r examineMergeClusters,eval=FALSE}
#re-does merging simpling to make plot 
#something like:
#assay(ceObj)<- ...
# if that replacement data should be considered on the transformed scale in plots, etc, the transformation function should be fixed as well:
#transformation(ceObj)<- ...
ceObj<-mergeClusters(ceObj, mergeMethod = "locfdr",
              plotInfo = "mergeMethod", cutoff = 0.01)
```

@@

EP: Fanny, you would need to use the argument 'clusterLegend'. That argument takes either the format of aheatmap (list with each element a *named* vector of colors) or the format of the clusterExperiment object (i.e. list with each element a matrix with columns for `name` and `color`). So I think the following code will run, though it might need the list to have names...

```{r}
sampleData <- data.frame(ours = primaryCluster(ceObj))
plotHeatmap(assays(se)$normalizedValues,
            main = 'Normalized values, 1000 most variable genes',
            clusterSamplesData = ceObj@dendro_samples,
            sampleData = as.matrix(sampleData),clusterLegend=ceObj@clusterLegend[1])
```

@@

EP: The function plotClusters aligns the clusters and can return the aligned coloring between clusterings. There is an option to return an updated CE object that changes the colors in clusterLegend to match the new colors found by aligning the clusters (`resetNames` and `resetColors` arguments). These haven't been extensively tested in real life because I don't think we've used them much. But it would be good to demonstrate!
